---
title: "HW2"
author: "Chaoqun Zhou"
date: 2022-10-04
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
```


# Problem 1

## Data Import:

```{r}
raw = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() 

p1 = select(raw, line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>% 
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE"),
         entry = as.logical(entry))
```

## Describe the data:
```{r}
skimr::skim(p1)
```
This data set is about NYC Transit data. So far, I have import the data into R. And use `janitor::clean_names()` to clean up variables names. And then I use `select()` to keep the columns that I want. And then I covert the entry variable from character to logical variable. Now, this data set has 19 columns and 1868 rows. It includes variables: `r ls(p1)`.These data looks tidy enough but it can be better. 

## Answer questions:

```{r}
x = distinct(p1, line, station_name)
```

1. There are 465 distinct stations. 


# Problem 2

## Data Import:

```{r}
tw = read_excel("./data/Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = ceiling(sports_balls), 
         sports_balls = as.integer(sports_balls))
```















