---
title: "HW2"
author: "Chaoqun Zhou"
date: 2022-10-04
output: github_document
---
```{r, message = FALSE}
library(tidyverse)
library(readxl)
library(stringr)
```


# Problem 1

## Data Import:

```{r, message = FALSE}
raw = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() 

p1 = select(raw, line, station_name, station_latitude, station_longitude,
    route1:route11, entrance_type, entry, vending, ada) %>% 
    mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE"),
         entry = as.logical(entry)) 
```

## Describe the data:
This data set is about NYC Transit data. So far, I have import the data into R. And use `janitor::clean_names()` to clean up variables names. And then I use `select()` to keep the columns that I want. And then I covert the entry variable from character to logical variable. Now, this data set has 19 columns and 1868 rows. It includes variables: `r ls(p1)`.These data looks tidy enough but it can be better. 

## Answer questions:

```{r, message = FALSE}
x = distinct(p1, line, station_name, .keep_all = TRUE)
mean(x$ada == 1)
mean(x$vending == "NO")
```

1. There are 465 distinct stations. 
2. There are `r mean(x$ada == 1)` stations that are ADA compliant.
3. There are `r mean(x$vending == "NO")` Stations that entrances / exits without vending.

## Reformat data:
```{r}
re = mutate(p1, route8 = as.character(route8),
            route9 = as.character(route9),
            route10 = as.character(route10),
            route11 = as.character(route11)) %>% 
pivot_longer(route1:route11, names_to = "route_number", values_to = "route_name")
```

```{r, message = FALSE}
y = distinct(re, line, station_name, .keep_all = TRUE)
mean(y$route_name == "A")
z = filter(y, route_name == c("A"))
mean(z$ada == 1)
```


# Problem 2

## Data Import:

Mr. Trash Wheel:
```{r, message = FALSE}
mrtw = read_excel("./data/Trash Wheel Collection Data.xlsx", 
                  sheet = "Mr. Trash Wheel", range = "A2:N549") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = ceiling(sports_balls), 
         sports_balls = as.integer(sports_balls)) %>% 
  mutate(which = "mrtw")
```

Professor Trash Wheel:
```{r, message = FALSE}
proftw = read_excel("./data/Trash Wheel Collection Data.xlsx", 
                    sheet = "Professor Trash Wheel", range = "A2:M96") %>% 
  janitor::clean_names() %>% 
  mutate(which = "proftw")
```

Combine two data set:
```{r, message = FALSE}
tw = merge(mrtw, proftw, all = TRUE)
```

Two Trash Wheels data set were imported. And then combined into one data set called `tw`. The final data set `tw` has `r nrow(tw)`observations. And it includes variable: `r ls(tw)`. The total weight of trash collected by Professor Trash Wheel `r sum(proftw$weight_tons)` tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(mrtw$sports_balls[mrtw$year == 2020])` tons.


# Problem 3

## Data Import and clean:

Clean the data in pols-month:
```{r, message = FALSE}
pols = read_csv(file = "./data/pols-month.csv") %>% 
    janitor::clean_names() %>% 
    separate(mon, c("year", "month", "day"), convert = TRUE) %>% 
    mutate(month = month.abb[month]) %>% 
    mutate(president = ifelse(prez_dem == 0, "gop", "dem")) %>% 
    select(-prez_dem, -prez_gop, -day)
```

Clean the data in SNP:
```{r, message = FALSE}
snp = read_csv(file = "./data/snp.csv") %>% 
    janitor::clean_names() %>% 
    separate(date, c("month", "day", "year"), convert = TRUE) %>%
    arrange(month) %>% 
    mutate(month = month.abb[month]) %>%
    mutate(year = ifelse(50 <= year & year <= 99, year + 1900, year + 2000)) %>%
    arrange(year) %>% 
    select(-day) %>% 
    relocate(year, month)
```

Clean the date in unemployment:
```{r, message = FALSE}
un = read_csv(file = "./data/unemployment.csv") %>% 
    janitor::clean_names() %>% 
    pivot_longer(jan:dec, names_to = "month", values_to = "percentage of unemployment") %>% 
    mutate(month = str_to_title(month))
```

Merging SNP into pols:
```{r, message = FALSE}
s1 = merge(snp, pols, all = TRUE)
```

Merging employment into result:
```{r, message = FALSE}
final = merge(un, s1, all = TRUE) %>%
  mutate(month = recode(month, Jan = "1", Feb = "2", Mar = "3", Apr = "4", 
                        May = "5", Jun = "6", Jul = "7", Aug = "8", Sep = "9", 
                        Oct = "10" ,Nov = "11", Dec = "12"), month = 
           as.integer(month)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.abb[month])
```

## Summary:

The data set `pols` is about the number of national politicians who are democratic or republican at any given time. It has `r nrow(pols)` rows and `r ncol(pols)` columns. Variables include: `r ls(pols)`. The year ranges is (`r range(pols$year)`).

The data set `snp` is about the Standard & Poor’s stock market index. It has `r nrow(snp)` rows and `r ncol(snp)` columns. Variables include: `r ls(snp)`. The year ranges is (`r range(snp$year)`).

The data set `un` is about the percentage of unemployment at any given time. It has `r nrow(un)` rows and `r ncol(un)` columns. Variables include: `r ls(un)`. The year ranges is (`r range(un$year)`).

The result data set `final` is generated by merging all three data sets mentioned above. We now have information on political parties of president, stock market index, and unemployment in one single data set. This data set has `r nrow(final)` rows and `r ncol(final)` columns. Variables include: `r ls(final)`. The year ranges is (`r range(final$year)`).










